### 2017-03-09 readme.Rmd
update to readme

p1_cell <- p1*(res(preds)[1]*res(preds)[2])

par(mfrow=c(1,3))

plot(p1_cell)
POdata <- species[species$Occurrence == 1,]
bkpts_grid <- generate_background_points(known_sites = POdata@coords,
                                    study_area = preds[[1]],
                                    model_covariates = preds,
                                    resolution = 2110, # this needs to be relative to raster resolution - this is about half the input resolution of the raster.
                                    method = 'grid')

@@

No let's check out estimates
```{r}
cellStats(p1_cell,sum)
cellStats(p2_cell,sum)

Let's compare the spatial prediction of the PPM against the PA species distribution model - hopefully we are in the right ball park
```{r,message=FALSE,error=FALSE,warning=FALSE}
d <- sdmData(formula= ~., train=species, predictors=preds)
dat <- cbind(species$Occurrence,d@features)
colnames(dat)[1]<-'occurrence'
fm3 <- gam(occurrence ~ s(elevation) +
              s(precipitation) +
              s(temperature) +
              s(vegetation),
              data = dat,
              family = binomial())

p3 <- predict(object=preds,
             model=fm3,
             type = 'response')
plot(p3)
```

The second function can be used to develop a bias layer which can be included as probabilities of including background points in the generation of quasirandom background points.

@@

@@

>>>>

p1_cell <- p1*(res(preds)[1]*res(preds)[2])
POdata <- species[species$Occurrence == 1,]
bkpts_grid <- generate_background_points(known_sites = POdata@coords,
                                    study_area = preds[[1]],
                                    model_covariates = preds,
                                    resolution = 2110, # this needs to be relative to raster resolution - this is about half the input resolution of the raster.
                                    method = 'grid')

@@

Now let's plot these models.
```{r,fig.width=12,fig.height=6}
jet.colors <- colorRampPalette(rev(RColorBrewer::brewer.pal(11 , "Spectral")))
par(mfrow=c(1,3),oma=c(4,4,4,4))
plot(p1_cell,col=jet.colors(100),main='quasi_random offset')
plot(p2_cell,col=jet.colors(100),main='grid offset')
plot(p_warton_cell,col=jet.colors(100),main='grid ala Warton2010')
```

Let's buld a PA species distribution based on the actuall occurrence data and see how we go.
```{r,message=FALSE,error=FALSE,warning=FALSE}
d <- sdmData(formula= ~., train=species, predictors=preds)
dat <- cbind(species$Occurrence,d@features)
colnames(dat)[1]<-'occurrence'
fm3 <- gam(occurrence ~ s(elevation) +
              s(precipitation) +
              s(temperature) +
              s(vegetation),
              data = dat,
              family = binomial())

p3 <- predict(object=preds,
             model=fm3,
             type = 'response')
```

Now let's compare the spatial predictions of the PPM against the PA species distribution model, hopefully we are in the right ball park.
```{r,fig.width=10,fig.height=8}
jet.colors <- colorRampPalette(rev(RColorBrewer::brewer.pal(11 , "Spectral")))
par(mfrow=c(2,2),oma=c(4,4,4,4))
plot(p1_cell,col=jet.colors(100),main='quasi_random offset')
plot(p2_cell,col=jet.colors(100),main='grid offset')
plot(p_warton_cell,col=jet.colors(100),main='grid ala Warton2010')
plot(p3,col=jet.colors(100),main='Bernoulli PA gam')
```

@@

In section 2.5 *Identifiability and sampling bias* from Fithian & Hastie (2014), they discuss the challanging concept of accounting for sampling bias in presence-only models. The discuss the idea of obtaining the *observation process*, which is difficult to do, as prence-only data only really contains information on the *sighting process* (absences aren't explicitly records and we only have information on sightings).

@@

Fithian, William, and Trevor Hastie. "Local case-control sampling: Efficient subsampling in imbalanced data sets." Annals of statistics 42.5 (2014): 1693.
